---
title: Analysis of Cognitive Decline Related Indicators
author: Yuyao Zhang
date: "`r Sys.Date()`"
output: html_document
---

```{r, here-i-am, include = FALSE}
here::i_am(
  "report.Rmd"
)
library(dplyr)
library(ggplot2)
library(kableExtra)
```

# Introduction

This analysis focuses on cognitive decline–related health indicators from the CDC Alzheimer's Disease and Healthy Aging Data. We examine percentages of older adults reporting subjective cognitive decline (SCD) and its impacts, using data from 2015–2022 across U.S. states. Our objective is to summarize these indicators and visualize their trends over time.

The dataset includes survey-based indicators of subjective cognitive decline (SCD) among older adults.
We focus on four specific survey questions directly measuring SCD and its consequences.

# Trend Over Time

```{r, echo=FALSE}
trend_data <- readRDS(
  here::here("output/trend_data.rds")
)

ggplot(trend_data, aes(x = Year, y = Avg_Percentage, color = Question_short)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Trends of Cognitive Decline Related Percentages Over Time",
    x = "Year",
    y = "Average Percentage (%)",
    color = "Indicator"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom")
```

The figure shows trends of subjective cognitive decline (SCD) indicators among older adults from 2015 to 2022. Overall, the percentages have remained relatively stable over time, with most indicators fluctuating only slightly. Indicators related to daily life interference and worsening memory symptoms consistently show higher prevalence (around 35–45%), while the indicator reflecting the need for assistance remains much lower (about 10–15%). This suggests that although many older adults experience cognitive decline symptoms, only a smaller proportion report needing help with day-to-day activities.

# Summary Table

```{r, echo=FALSE}
derived_data <- readRDS(
  here::here("output/derived_data.rds")
)

summary_table <- derived_data %>%
  group_by(Question_short) %>%
  summarise(
    Mean_Percentage = round(mean(Data_Value, na.rm = TRUE), 2),
    SD_Percentage = round(sd(Data_Value, na.rm = TRUE), 2),
    Observations = n(),
    .groups = "drop"
  )

kable(summary_table, caption = "Summary Statistics of Cognitive Decline Related Percentages") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  row_spec(0, bold = TRUE)
```

The findings suggest that while a significant portion of older adults experience cognitive decline that impacts daily functioning, relatively few report worsening symptoms over time and less than half seek professional help. This highlights a potential gap in awareness, monitoring, and health care engagement. The table provides summary statistics across different cognitive decline–related survey questions. For instance, the mean percentages highlight differences in prevalence across questions.
